Literature review of Formal Methods applied to Blockchain technology

### 2022-01-14 article/article.tex
finished article

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
  Vestibulum enim nibh, aliquet sit amet ex suscipit, efficitur fringilla enim. 
  Vestibulum elementum fringilla sodales. Cras malesuada sollicitudin consectetur.
  Cras elit nibh, tempus elementum dui vitae, sagittis porttitor elit. In sollicitudin purus neque, eu tristique felis varius in. Morbi sed commodo eros.
  Pellentesque ornare sed felis et bibendum.
\end{abstract}

@@

\section{Introduction}
The rising complexity of features offered by software gives us the ability to use it for the most integrity and reliability dependent needs of our lives. 
Information systems that rely on these base principles can be financial and health related services, and in most recent times, blockchain based applications. 
Blockchain, being a decentralized and distributed network protocol that can be used as the core of a monetary system that performs peer-to-peer transactions, 
it needs the assurance of coordination and consensus of its economy’s state~\cite{Duan2018}. 
Knowing that the protocol of a given blockchain project is what dictates how transactions and their coordination is maintained, the assurance of security and safety is a given. 
Formal methods can be introduced in order to provide unequivocal evidence that a given blockchain system and its consensus algorithm are secure and in according to expected software quality.

@@

This paper is organized as follows. Section \ref{formalModeling} introduces applications of formal modeling of blockchain technology.
Section \ref{consensusRule} expands on the use of formal mathematical descriptions of a blockchain's consensus rule.
Section \ref{smartContracts} explored how formal modeling can be used to verify and assure quality to smart contracts using Colored Petri Nets (CPN).
Section \ref{security} elaborates on the security requirements of a blockchain system.
Section \ref{conclusion} is a brief conclusion.

@@

  \begin{lstlisting}[frame=single,label={lst:block-code},
    caption={Definition of a block's structure},captionpos=b, basicstyle=\ttfamily]
    newtype blockchain struct
      prehash integer;
      hash integer;
      length integer;/*the length of blocks*/
      merkleroot integer;
      ti Time;/*Timestamp*/
      translist list;
    endnewtype;

@@

@@

\begin{displaymath}
  H_ = \left(\sum_{i}^{n} I(s_i.e \wedge (s_i.v = b)) \geq m \right)
\end{displaymath}

@@

  \begin{table}
    \caption{Blockchain Technology layers and security considerations}
    \label{tab:security-layers}
    \begin{tabularx}{\linewidth}{ccl}
      \toprule
      Layer&Security Consideration& Standard \\
      \toprule
      Operation & \makecell{Key Management, \\Audit, Backup} & ISO/IEC 27000 \\ 
      \midrule
      Implementation & \makecell{Program Code,\\Secure Hardware} & ISO/IEC 15408 \\ 
      \midrule
      Application Logic & \makecell{Scripting Language\\for Financial\\Transaction,\\ Contract} & \makecell{Secure coding\\guides} \\ 
      \midrule
      Application Protocol & \makecell{Privacy protection,\\Secure transaction} & ISO/IEC 29128 \\ 
      \midrule
      Backbone Protocol & \makecell{PSP, Consensus,\\Merkle Tree} & ISO/IEC 29128 \\ 
      \midrule
      Cryptography & \makecell{ECDSA, SHA-2,\\RIPEMD160} & NIST, ISO \\
      \bottomrule
  \end{tabularx}
  \end{table}

@@

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Blockchain is a technology that allows the decentralized record keeping and monetary transactions.
  Its different uses are constantly being explored and new projects emerge on a daily basis.
  The underlying design of such systems can bring about dire consequences if deployed with an unclear design
  and possible vulnerabilities. If we use formal methods to mathematically model the base structures of out blockchain
  system we provide irrefutable proof that our design will work as intended. This includes the usage of modeling languages
  to describe our system and block, mathematical descriptions of the applied consensus rule and endorsement policy and its different
  security requirements, including a smart contract that contains the code of an application hosted in a blockchain.

\end{abstract}

@@

\section{Introduction}
The rising complexity of features offered by software gives us the ability to use it for the most integrity and reliability dependent needs of our lives. 
Information systems that rely on these base principles can be financial and health related services, and in most recent times, blockchain based applications. 
Blockchain, being a decentralized and distributed network protocol that can be used as the core of a monetary system that performs peer-to-peer transactions, 
it needs the assurance of coordination and consensus of its economy’s state~\cite{Duan2018}. 
Knowing that the protocol of a given blockchain project is what dictates how transactions and their coordination is maintained, the assurance of security and safety is a given. 
Formal methods can be introduced in order to provide unequivocal evidence that a given blockchain system and its consensus algorithm and inherent smart contracts are secure and in according to expected software quality.

@@

This paper is organized as follows. Section~\ref{formalModeling} introduces applications of formal modeling of blockchain technology.
Section~\ref{consensusRule} expands on the use of formal mathematical descriptions of a blockchain's consensus rule.
Section~\ref{smartContracts} explored how formal modeling can be used to verify and assure quality to smart contracts using Colored Petri Nets (CPN).
Section~\ref{security} elaborates on the security requirements of a blockchain system.
Section~\ref{conclusions} is a brief conclusion.

@@

  \begin{lstlisting}[frame=single,label={lst:block-code},
    caption={Definition of a block's structure},captionpos=b, basicstyle=\ttfamily\small]
    newtype blockchain struct
      prehash integer;
      hash integer;
      length integer;/*the length of blocks*/
      merkleroot integer;
      ti Time;/*Timestamp*/
      translist list;
    endnewtype;

@@

  An example of a diagram that describes the system and a block can be seen on Figure~\ref{fig:system-block}.

  \begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{system-block-diagram}
    \caption{System and block structure diagram, from~\cite{Duan2018}}
    \Description{Ilustration of the layers that describe a blockchain system}
    \label{fig:system-block}
  \end{figure}

@@

\begin{displaymath}
  H_b = \left(\sum_{i}^{n} I(s_i.e \wedge (s_i.v = b)) \geq m \right)
\end{displaymath}

@@

  \begin{table}
    \caption{Blockchain Technology layers and security considerations}
    \label{tab:security-layers}
    \begin{tabularx}{\linewidth}{ccl}
      \toprule
      Layer&Security Consideration& Standard \\
      \toprule
      Operation & \makecell{Key Management,\\Audit, Backup} & ISO/IEC 27000 \\ 
      \midrule
      Implementation & \makecell{Program Code,\\Secure Hardware} & ISO/IEC 15408 \\ 
      \midrule
      Application Logic & \makecell{Scripting Language\\for Financial\\Transaction,\\ Contract} & \makecell{Secure coding\\guides} \\ 
      \midrule
      Application Protocol & \makecell{Privacy protection,\\Secure transaction} & ISO/IEC 29128 \\ 
      \midrule
      Backbone Protocol & \makecell{PSP, Consensus,\\Merkle Tree} & ISO/IEC 29128 \\ 
      \midrule
      Cryptography & \makecell{ECDSA, SHA-2,\\RIPEMD160} & NIST, ISO \\
      \bottomrule
  \end{tabularx}
  \end{table}

@@

\section{Conclusions}\label{conclusions}
If we apply formal methods to Blockchain technology and its inherent structures and properties we can have a clear vision of the implementation
of its design.

The nodes of a blockchain system can be mathematically described in a way that we can predict its behavior.
The same can be done for the state of the blockchain at a given moment. The same mathematical predictability
can be applied to the description of endorsement policies. Smart contracts also benefit from formal verification,
since their code can be an application used by common users, and CPN with tools like Space State Tools can provide
a way of formally verifying the design of a contract.
A blockchain system that was planned using formal methods benefit from assured security in the requirement layers
that were considered.
